---
title: "Integration"
editor_options: 
  chunk_output_type: console
---


## Material

{{< downloadthis ../assets/pdf/08_Integration.pdf dname="08_Integration" label="Download the presentation" icon="filetype-pdf" >}}

{{< video https://youtu.be/2TsW5A53hTg?si=Bas__YskLl21-YjF >}}

## Exercises

```{r day2-3-integration-1}
#| warning: false
#| message: false

library(Seurat)
seu <- readRDS("day2/seu_day2-2.rds")
```

Let's have a look at the UMAP again. Although cells of different samples are shared amongst 'clusters', you can still see seperation within the clusters:

```{r day2-3-integration-2}
Seurat::DimPlot(seu, reduction = "umap")
```

Let's also make the function again for 3D UMAP

```{r day2-3-integration-3, warning=FALSE, message=FALSE}
library(plotly)
seu <- Seurat::RunUMAP(seu, dims = 1:25, n.components = 3, reduction.name = "umap_3D")

plot_3d_umap <- function(seu,
                         reduction = "umap_3D",
                         sample_col = "orig.ident",
                         point_size = 4,
                         point_opacity = 0.5) {
  
  ## 1. Extract coordinates -------------------------------------------------
  umap_coords <- as.data.frame(Embeddings(seu, reduction = reduction))
  umap_coords$Sample <- seu@meta.data[[sample_col]]
  umap_coords$Cell   <- rownames(umap_coords)   # for hover
  
  ## 2. Build Plotly object --------------------------------------------------
  p <- plot_ly(
    data = umap_coords,
    x = ~umap3D_1, y = ~umap3D_2, z = ~umap3D_3,
    color = ~Sample,                     # default Plotly palette
    type = "scatter3d",
    mode = "markers",
    marker = list(
      size = point_size,
      opacity = point_opacity
    ),
    text = ~paste("Cell:", Cell, "<br>Sample:", Sample),
    hovertemplate = "<b>%{text}</b><extra></extra>"
  ) %>%
    layout(
      title = list(
        text = "<b>3D UMAP of Cells by Sample</b>",
        font = list(size = 18, family = "Arial, sans-serif"),
        x = 0.5, xanchor = "center", y = 0.95
      ),
      scene = list(
        xaxis = list(
          title = "<b>UMAP 1</b>", titlefont = list(size = 14),
          showgrid = TRUE, gridcolor = "rgba(200,200,200,0.5)",
          backgroundcolor = "rgba(245,245,255,0.9)",
          zerolinecolor = "rgba(100,100,100,0.3)"
        ),
        yaxis = list(
          title = "<b>UMAP 2</b>", titlefont = list(size = 14),
          showgrid = TRUE, gridcolor = "rgba(200,200,200,0.5)",
          backgroundcolor = "rgba(245,245,255,0.9)",
          zerolinecolor = "rgba(100,100,100,0.3)"
        ),
        zaxis = list(
          title = "<b>UMAP 3</b>", titlefont = list(size = 14),
          showgrid = TRUE, gridcolor = "rgba(200,200,200,0.5)",
          backgroundcolor = "rgba(245,245,255,0.9)",
          zerolinecolor = "rgba(100,100,100,0.3)"
        ),
        camera = list(eye = list(x = 1.5, y = 1.5, z = 1.5)),
        aspectmode = "cube"
      ),
      legend = list(
        title = list(text = "<b>Sample</b>", font = list(size = 12)),
        bgcolor = "rgba(255,255,255,0.8)",
        bordercolor = "gray", borderwidth = 1
      ),
      margin = list(l = 0, r = 0, t = 60, b = 0),
      paper_bgcolor = "rgba(250,250,252,1)",
      plot_bgcolor  = "rgba(250,250,252,1)"
    ) %>%
    config(
      displayModeBar = TRUE,
      displaylogo = FALSE,
      modeBarButtonsToRemove = c("sendDataToCloud", "lasso2d", "select2d")
    )
  
  return(p)
}

plot_3d_umap(seu = seu, reduction = "umap_3D", sample_col = "orig.ident")
```


To perform the integration, we split our object by sample, resulting into a set of layers within the `RNA` assay. The layers are integrated and stored in the reduction slot - in our case we call it `integrated.cca`. Then, we re-join the layers

::: {.panel-tabset}

## RNA Assay

```{r day2-3-integration-4}
#| output: false

seu[["RNA"]] <- split(seu[["RNA"]], f = seu$orig.ident)

seu <- Seurat::IntegrateLayers(object = seu, method = CCAIntegration,
                       orig.reduction = "pca",
                       new.reduction = "integrated.cca",
                       verbose = FALSE)

# re-join layers after integration
seu[["RNA"]] <- JoinLayers(seu[["RNA"]])
```

## SCT Assay
```{r day2-3-integration-5}
# 1. Ensure default assay & have your HVGs / PCA ready
DefaultAssay(seu) <- "SCT"

# If pcaSCT is missing or outdated:
seu <- RunPCA(seu, reduction.name = "pcaSCT", verbose = FALSE)

# 2. Create a new v5 assay from your current SCT (copies data/counts/scale.data)
#    → SCT residuals go into 'data' layer; this becomes Assay5 class
seu[["SCT_v5"]] <- as(object = seu[["SCT"]], Class = "Assay5")

# Confirm class & layers
class(seu[["SCT_v5"]])          # should be "Assay5"
Layers(seu[["SCT_v5"]])         # expect "counts", "data", possibly "scale.data"

# Optional: If you want split layers (recommended for integration):
seu[["SCT_v5"]] <- split(seu[["SCT_v5"]], f = seu$orig.ident)
Layers(seu[["SCT_v5"]])         # now per-batch layers (counts.PBMMC-1, data.PBMMC-1, etc.)

# 3. Run CCAIntegration on the v5 assay
seu <- IntegrateLayers(
  object         = seu,
  assay          = "SCT_v5",                 # ← use the new v5 assay here
  method         = CCAIntegration,
  orig.reduction = "pcaSCT",                 # your existing PCA (computed on SCT)
  new.reduction  = "SCT.cca",
  features       = VariableFeatures(seu),    # still good to restrict
  verbose        = FALSE
)

DefaultAssay(seu) <- "RNA"
```

:::

We can then use this new integrated matrix for clustering and visualization. Now, we can re-run and visualize the results with UMAP.

::: {.callout-important}
## Exercise
Create the UMAP again on the `integrated.cca` reduction (using the function `RunUMAP` - set the option `reduction` accordingly). After that, generate the UMAP plot. Did the integration perform well?
::: 

::: {.callout-tip collapse="true"}
## Answer

Performing the scaling, PCA and UMAP:

```{r day2-3-integration-6}
#| output: false
seu <- RunUMAP(seu, dims = 1:25, reduction = "integrated.cca")
```

Plotting the UMAP:

```{r day2-3-integration-7}
Seurat::DimPlot(seu, reduction = "umap")
```
::: 


**Also in 3D**
```{r day2-3-integration-8}
seu <- RunUMAP(seu, dims = 1:25, n.components = 3, , reduction = "integrated.cca", reduction.name = "umap_3D")

plot_3d_umap(seu = seu, reduction = "umap_3D", sample_col = "orig.ident")
```


### Save the dataset and clear environment


```{r day2-3-integration-9}
saveRDS(seu, "day2/seu_day2-3.rds")
```

Clear your environment:

```{r day2-3-integration-10}
#| eval: false
gc()
```

